PACMAN deployment summary
========================================================
Load data
-------------------------------
The data from PACMAN does not include an ISO compliant date field so it must be constructed from the available data
```{r prepare_environment}
library('openair')
library('ggplot2')
unitID<-'pacman_18'
pacman.data <- read.delim(paste0("/home/gustavo/data/CONA/PACMAN/campaign/",unitID,".txt"))
names(pacman.data)<-c('Count','Year','Month','Day','Hour','Minute','Second',
                      'Distance','Temperature_IN_C','Temperature_mV','PM_mV','CO2_mV','CO_mV','Movement','COstatus')
pacman.data$Temperature_mV <- pacman.data$Temperature_mV/1000
pacman.data$date<-ISOdatetime(year=pacman.data$Year,
                              month=pacman.data$Month,
                              day=pacman.data$Day,
                              hour=pacman.data$Hour,
                              min=pacman.data$Minute,
                              sec=pacman.data$Second,
                              tz="GMT")
dstart=as.character(min(pacman.data$date))
dend=as.character(max(pacman.data$date))
```

Deployment details
--------------------------------
Parameter | Value
------------- | -------------
Date start | `r dstart`
Date finish | `r dend`
Location | NIWA Christchurch
Unit | `r unitID`


Dust correction
------------------------------
Using data from previous deployments of similar dust sensors (ODIN project), the raw voltage reported by PACMAN was corrected to compensate for the impact of ambient temperature on its operation.

```{r correct_Dust}
## Dust = dustA*PM_mV + dustB*Temperature_mV + dustC ug/m3 ##

# dustA=0.8798
# dustB=-3.5372
# dustC=-289.6213
# pacman.data$Dust<-dustA*pacman.data$PM_mV + dustB*pacman.data$Temperature_mV + dustC
pacman.data$Dust<-pacman.data$PM_mV
```

Carbon Monoxide data
------------------------------
The sensor used by PACMAN operates in cycles and only one data point per cycle is valid. PACMAN reports the `CO_status` flag that indicates the position in the cycle that the sensor is at. The valid data point correspond to the end of the `CO_status` labelled 2.

```{r co_label}
nrecs=length(pacman.data$COstatus)
pacman.data$CO_ok<-FALSE
pacman.data$CO_ok[-1]=(pacman.data$COstatus[-1]==1)
```

Weekly variation plots
-----------------------------
To summarize the date, we produced weekly diurnal variation plots for the measurements.

### Distance
```{r distance,fig.keep='none'}
diurnal.distance<-timeVariation(pacman.data,pollutant='Distance')
```
```{r distance_plot}
ggplot(diurnal.distance$data$day.hour)+
  geom_ribbon(aes(x=hour,ymin=Lower,ymax=Upper),fill='red')+
  geom_line(aes(x=hour,y=Mean),colour='red')+
  facet_grid(~wkday)+
  ggtitle('Distance')+
  xlab('NZST hour')+
  ylab('Distance from sensor [mV]')
```

### Temperature outside
```{r temp_out,fig.keep='none'}
diurnal.temp_out<-try(timeVariation(pacman.data,pollutant='Temperature_IN_C'))
```
```{r temp_out_plot}
if (class(diurnal.temp_out)!="try-error"){
  ggplot(diurnal.temp_out$data$day.hour)+
    geom_ribbon(aes(x=hour,ymin=Lower,ymax=Upper),fill='red')+
    geom_line(aes(x=hour,y=Mean),colour='red')+
    facet_grid(~wkday)+
    ggtitle('External temperature')+
    xlab('NZST hour')+
    ylab('External Temperature [C]')
}
```

### Temperature inside
```{r temp_in,fig.keep='none'}
diurnal.temp_in<-try(timeVariation(pacman.data,pollutant='Temperature_mV'))
```
```{r temp_in_plot}
if (class(diurnal.temp_in)!="try-error"){
  ggplot(diurnal.temp_in$data$day.hour)+
    geom_ribbon(aes(x=hour,ymin=Lower,ymax=Upper),fill='red')+
    geom_line(aes(x=hour,y=Mean),colour='red')+
    facet_grid(~wkday)+
    ggtitle('Internal Temperature')+
    xlab('NZST hour')+
    ylab('Internal Temperature [C]')
}
```

### Dust
```{r dust,fig.keep='none'}
diurnal.dust<-try(timeVariation(pacman.data,pollutant='Dust'))
```
```{r dust_plot}
if (class(diurnal.dust)!="try-error"){
ggplot(diurnal.dust$data$day.hour)+
  geom_ribbon(aes(x=hour,ymin=Lower,ymax=Upper),fill='red')+
  geom_line(aes(x=hour,y=Mean),colour='red')+
  facet_grid(~wkday)+
  ggtitle('Dust \n Corrected with ODIN data')+
  xlab('NZST hour')+
  ylab('Dust [ug/m3]')
}
scatterPlot(pacman.data,x='Temperature_mV',y='Dust')
```

### Carbon Dioxide
```{r co2,fig.keep='none'}
diurnal.co2<-try(timeVariation(pacman.data,pollutant='CO2_mV'))
```
```{r co2_plot}
if (class(diurnal.co2)!="try-error"){
ggplot(diurnal.co2$data$day.hour)+
  geom_ribbon(aes(x=hour,ymin=Lower,ymax=Upper),fill='red')+
  geom_line(aes(x=hour,y=Mean),colour='red')+
  facet_grid(~wkday)+
  ggtitle('CO2')+
  xlab('NZST hour')+
  ylab('CO2 [mV]')
}
```

### Carbon Monoxide
```{r co,fig.keep='none'}
diurnal.co<-try(timeVariation(pacman.data,pollutant='CO_mV',plot=FALSE))
```
```{r co_plot}
if (class(diurnal.co)!="try-error"){
ggplot(diurnal.co$data$day.hour)+
  geom_ribbon(aes(x=hour,ymin=Lower,ymax=Upper),fill='red')+
  geom_line(aes(x=hour,y=Mean),colour='red')+
  facet_grid(~wkday)+
  ggtitle('CO')+
  xlab('NZST hour')+
  ylab('CO [mV]')
}
```

### Movement
```{r move,fig.keep='none'}
diurnal.mov<-try(timeVariation(pacman.data,pollutant='Movement'))
```
```{r move_plot}
if (class(diurnal.mov)!="try-error"){
ggplot(diurnal.mov$data$day.hour)+
  geom_ribbon(aes(x=hour,ymin=Lower,ymax=Upper),fill='red')+
  geom_line(aes(x=hour,y=Mean),colour='red')+
  facet_grid(~wkday)+
  ggtitle('Movement')+
  xlab('NZST hour')+
  ylab('Movement')
  }
```
