---
title: "IButton event detection"
output: html_notebook
---
# Abstract

The main objective of the deployment of IButton temperature sensors in the [CONA](http://www.niwa.co.nz/cona) project was to obtain independent information about when the participants used their wood burners. We present here a test using [Twitter's](www.twitter.com) **Anomaly Detection** package to identify these wood burning events. For the 2016 campaign we have two weeks of data for 8 homes. Each home had two IButtons installed, one on or very near the hot part of the wood burner (flue, firebox, door) and a second one in the main living area away from the wood burner.
We found that ....

## Required libraries and paths
```{r load_libraries}
library(RPostgreSQL)
library(openair)
library(ggplot2)
library(reshape2)
library(AnomalyDetection)
library(readr)
working_path <- "~/repositories/cona/"
data_path <- "~/data/CONA/2016/iButton/"
```

# Data

For this test we use data from subject Nr 3 who received 2 iButtons (SN1: A500000032322841; SN2: B00000026F94E41).



```{r load_data_ibuttons}
ibutton.1 <- read_csv("~/data/CONA/2016/iButton/A0000002F194D41.csv",
                      col_types = cols(Unit = col_skip()),
                      skip = 19)
ibutton.1$date <- as.POSIXct(ibutton.1$`Date/Time`, tz='NZST', format = '%d/%m/%y %I:%M:%S %p')
ibutton.1$`Date/Time`<- NULL
names(ibutton.1) <- c('T.1','date')
ibutton.2 <- read_csv("~/data/CONA/2016/iButton/170000004AE5E141.csv",
                      col_types = cols(Unit = col_skip()),
                      skip = 19)
ibutton.2$date <- as.POSIXct(ibutton.2$`Date/Time`, tz='NZST', format = '%d/%m/%y %I:%M:%S %p')
ibutton.2$`Date/Time` <- NULL
names(ibutton.2) <- c('T.2','date')

# Merge data
ibutton <- merge(ibutton.1,ibutton.2,by = 'date', all = TRUE)
```

# Analysis
## IButton comparison
Comparing the range of temperatures of both devices it is clear that ibutton1 was located on the wood burner while ibutton2 was on the living area. It is also worth noting that ibutton1 reached both its lower and upper temperature limits which could affect the event detection algorithm.
```{r temperature_compare, fig.height=6, fig.width=12}
summary(ibutton[,2:3])
timePlot(ibutton,pollutant = c('T.1', 'T.2'), avg.time = '5 min', group = TRUE)
timeVariation(ibutton,pollutant = c('T.1', 'T.2'))
```
## Anomaly Detection
To homogenise the timebase of the ibutton data we created a 5m average time series

```{r aggregate_5_min}
ibutton.5min <- timeAverage(ibutton,avg.time = '5 min')
```

Now, let's see what the **Anomaly Detection** library produces when fed with the ibutton data we have.

### IButton 1
```{r ibutton1_anomaly, fig.height=6, fig.width=12}
anomalies.T1 <- AnomalyDetectionTs(ibutton.5min[,c('date','T.1')],
                                   max_anoms = 0.3,
                                   plot = TRUE,
                                   title = 'IButton 1',
                                   alpha = 0.3,
                                   direction = 'both')
```

The algorithm is able to capture most, but not all, instances of high temperature which indicate operation of the wood burner. Changing some of the parameters the performance decreases significantly. For example, choosing a shorter period for the data, the algorithm only identifies the temperature saturation events and changing the significance criteria and the maximum percentage of anomalies changes the number of events identified.

### IButton 2
```{r ibutton2_anomaly, fig.height=6, fig.width=12}
anomalies.T2 <- AnomalyDetectionTs(ibutton.5min[,c('date','T.2')],
                                   max_anoms = 0.3,
                                   plot = TRUE,
                                   title = 'IButton 2',
                                   alpha = 0.3,
                                   direction = 'both')
```


### What about the difference between sensors?
In an ideal setup, the *hot* sensor (wood burner) would only deviate from the *cold* sensor (living room) when the woodburner is in use and that might give us another way to identify when the woodburner was operational.
```{r difference_temperature,  fig.height=6, fig.width=12}
ibutton.5min$DT <- ibutton.5min$T.1 - ibutton.5min$T.2
ibutton.5min$RT <- ibutton.5min$T.1 / ibutton.5min$T.2
timePlot(ibutton.5min,pollutant = c('DT','RT'))

```
Now let's extract the anomalies from these new columns
```{r anomalies_ratios}
anomalies.delta <- AnomalyDetectionTs(ibutton.5min[,c('date','DT')],
                   max_anoms = 0.3,
                   plot = TRUE,
                   title = 'T1 - T2 ',
                   alpha = 0.3,
                   direction = 'pos')
anomalies.ratio <- AnomalyDetectionTs(ibutton.5min[,c('date','RT')],
                   max_anoms = 0.3,
                   plot = TRUE,
                   title = 'T1 / T2',
                   alpha = 0.3,
                   direction = 'pos')
```
How different are these anomalies?
```{r plot_anomalies, fig.height=6, fig.width=12}
anomalies.T1$anoms$event <- anomalies.T1$anoms$anoms / max(anomalies.T1$anoms$anoms)
anomalies.T2$anoms$event <- anomalies.T2$anoms$anoms / max(anomalies.T2$anoms$anoms)
anomalies.delta$anoms$event <- anomalies.delta$anoms$anoms / max(anomalies.delta$anoms$anoms)
anomalies.ratio$anoms$event <- anomalies.ratio$anoms$anoms / max(anomalies.ratio$anoms$anoms)
plot.t1 <- geom_point(aes(x=anomalies.T1$anoms$timestamp,y=anomalies.T1$anoms$event), color = 'black')
plot.t2 <- geom_point(aes(x=anomalies.T2$anoms$timestamp,y=anomalies.T2$anoms$event), color = 'red')
plot.dt <- geom_point(aes(x=anomalies.delta$anoms$timestamp,y=anomalies.delta$anoms$event), color = 'green')
plot.rt <- geom_point(aes(x=anomalies.ratio$anoms$timestamp,y=anomalies.ratio$anoms$event), color = 'blue')
plot.t1.t2 <- geom_point(aes(x=anomalies.T1$anoms$timestamp,y=anomalies.T1$anoms$event), color = 'black')
p <- ggplot()
p +
  plot.t1 +
  plot.t2 +
  plot.dt +
  plot.rt

```

Let's explore the whether the anomalies identified on these four time series correspond in time.

```{r merge_anomalies}

```



